Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-05-27 WSSSPE2_report/WSSSPE2_review.tex
more work in 10

Thaney spoke about Mozilla's work around code citation, through a collaboration
and prototype crafted between Mozilla, GitHub, figshare, and Zenodo. This work
was presented at a closed meeting in May 2014 at the National Institutes of
Health (NIH) around these issues, sparking a conversation from that meeting
around what a \emph{Software Discovery
Index}\footnote{Software Discovery Index, \url{http://softwarediscoveryindex.org/report/}} might look
like. The meeting included a number of publishers, researchers, and those behind
major scientific software efforts such as 
Bioconductor{\footnote{Bioconductor, \url{http://www.bioconductor.org}}, 
Galaxy\footnote{Galaxy, \url{http://galaxyproject.org}}, and
nanoHUB\footnote{nanoHUB, \url{https://nanohub.org}}. 
%facilitates more efficient scientific research. SDI identifies scientific
%software by archiving and standardizing metadata for software and hence help
%connect both developers and users.
Ted Habermann in the audience commented that if the metadata is minimal, it
would be less onerous for data providers, but more burdensome for users---it
could be challenging to keep a balance between what have to be captured and what
would be ideal if we do not want to lose user engagement as in the case of the
old Harvard Dataverse, finding often only the first four fields of three pages
of metadata were filled out.

@@

\todo{short intro to group here}

@@

This group began by recognizing that they know what the code testing and code review problems are, so the purpose of the group was really to think about concrete things that people in the group would be willing to commit to, in order to improve the sustainability of software.  It was suggested that these concrete things should be small things so that the group members were likely to do them.

@@

in progress....

%Paper on multiphysics software verificationt01 - \url{http://onlinelibrary.wiley.com/doi/10.1002/spe.2220/abstract}
%
%Don't accept patches that don't have tests
%
%Setting up infrastructure to test well is difficult
%
%Want a standard way that runs out of the box
%Q: what about Jenkins?
%A: sure, but you have to set it up, and that's not easy for most scientists
%
%People who've worked with a testing framework once (even if they have to be forced to do so) won't work without it afterward
%Q: but how to get them to do it the first time?
%
%Two important pieces of the barrier:
%a. Picking/implementing testing frameworks (technical barrier)
%b. What are the actual tests that I write?
%Ask students ``how is what you're doing different from typing random keys on the keyboard?'' Turn their answer into the concept of tests.
%
%How many scientists are aware of the fact that ``testing the code'' and ``testing the science'' are separate?
%
%``Every scientist should go through a programming class''
%a. That's probably too big a goal for this group for the next year
%b. Software Carpentry exists because university curricula are full; can't add extra classes
%
%
%Validation questions: how do you know that your method produces a result that is ‘close enough’
%
%Tom Clune wants a piece of code that badly engineered/lacks unit tests so that he can add the unit tests and show people what it should look like
%a. The collaboration must be with someone that either wrote the code or is intimately familiar with the topic.  The exercise will be to extract unit-tests through reverse-engineering {\b and} questioning the developer/expert
%b. Code does not have to be ``badly engineered'', but is either representative or is believed to be impenetrable from a unit-testing perspective.

@@

Useful reading:
Daniel Hook, Diane Kelly, "Mutation Sensitivity Testing", Computing in Science \& Engineering, vol.11, no. 6, pp. 40-47, November/December 2009, doi:10.1109/MCSE.2009.200
Daniel Hook, Diane Kelly, "Testing for trustworthiness in scientific software", ICSE Workshop on Software Engineering for Computational Science and Engineering (SECSE 2009), 2009 pp. 59-64, doi:10.1109/SECSE.2009.5069163

See the “13 dwarves” from Berkeley http://www.eecs.berkeley.edu/Pubs/TechRpts/2006/EECS-2006-183.html

>>>>

Thaney spoke about Mozilla's work around code citation, through a collaboration
and prototype crafted between Mozilla, GitHub, figshare, and Zenodo. This work
was presented at a closed meeting in May 2014 at the National Institutes of
Health (NIH) around these issues, sparking a conversation from that meeting
around what a \emph{Software Discovery
Index}\footnote{Software Discovery Index, \url{http://softwarediscoveryindex.org/report/}} might look
like. The meeting included a number of publishers, researchers, and those behind
major scientific software efforts such as 
Bioconductor\footnote{Bioconductor, \url{http://www.bioconductor.org}}, 
Galaxy\footnote{Galaxy, \url{http://galaxyproject.org}}, and
nanoHUB\footnote{nanoHUB, \url{https://nanohub.org}}. 
%facilitates more efficient scientific research. SDI identifies scientific
%software by archiving and standardizing metadata for software and hence help
%connect both developers and users.
Ted Habermann in the audience commented that if the metadata is minimal, it
would be less onerous for data providers, but more burdensome for users---it
could be challenging to keep a balance between what have to be captured and what
would be ideal if we do not want to lose user engagement as in the case of the
old Harvard Dataverse, finding often only the first four fields of three pages
of metadata were filled out.

@@

This group began by recognizing that they know what the code testing and code review problems are, so the purpose of the group was really to think about concrete things that people in the group would be willing to commit to, in order to improve the sustainability of software.

@@

It was suggested that these concrete things should be small things so that the group members were likely to do them.

@@

Some lessons on multiphysics software verification can be found in a recent paper~\cite{dubey-FLASH}.

One general rule that was suggested is that projects shouldn't accept patches that don't have associated tests.  This led to discussion about the fact that setting up infrastructure to test well is difficult.  The group would like a standard way that runs out of the box.  Jenkins was suggested as
a solution, but setting it up is considered difficult for most scientists.

Group members also comments that people who've worked with a testing framework once (even if they have to be forced to do so) won't work without it afterward.  But how to get them to do this the first time is not clear.

Two important pieces of the barrier are a) picking/implementing testing frameworks (a technical barrier) and b) deciding what are the actual tests that I need to write?  A possible method to solve the latter is to ask students ``how is what you're doing different from typing random keys on the keyboard?'' and then, turn their answer into the concept of tests.

Another point of discussion is that many scientists are not aware that ``testing the code'' and ``testing the science'' are separate.  A related question that came up is ``How do you know that your method produces a result that is `close enough'?''

Some group members thought that every scientist should go through a programming class, but that was considered too big a goal for the group for the next year.  Also, it's not clear where these classes could be offered.  Software Carpentry exists because university curricula are full; they can't add extra classes.

Tom Clune suggested it would be useful to have a piece of code that is badly engineered and lacks unit tests so that he can add the unit tests and show people what it should look like.  However, such a collaboration must be with someone that either wrote the code or is intimately familiar with the topic.  The exercise will be to extract unit-tests through reverse-engineering {\b and} questioning the developer/expert.
Others suggested that the code does not have to be ``badly engineered'', but is either representative or is believed to be impenetrable from a unit-testing perspective.

@@

The group suggested two papers by Hook and Kelly that would be useful for those interested in software testing to read~\cite{hook1,hook2}

See the ``13 dwarves'' from Berkeley http://www.eecs.berkeley.edu/Pubs/TechRpts/2006/EECS-2006-183.html \katznote{to delete? - don't see how it fits}

\subsection{Testing examples}

The group took some time to define

\katznote{... working from here still...}
